Direction-Selective Resistance to Cerebrospinal-Fluid Flow As the Mechanism of Syrinx Generation in Syringomyelia

### 2022-07-12 manuscript/main.tex
minor revision

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\title{Direction-Selective Resistance to Cerebrospinal-Fluid Flow As
the Mechanism of Syrinx Generation in Syringomyelia}

@@

\keywords{syringomyelia, pathophysiology, simulation}

\begin{abstract}
    The pathophysiology of syringomyelia is not well understood. The main
    theoretical problem is how cerebrospinal fluid (CSF) enters from
    low-pressure subarachnoid space to high-pressure syrinx and stays
    there.
\end{abstract} \begin{document}

@@

\noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.

@@

The pathophysiology of syringomyelia is still poorly understood. A number
of hypotheses exist in the literature \cite{gardner1958mechanism,
williams1980pathogenesis, milhorat1999chiari, ball1972pathogenesis,
klekamp2002pathophysiology, duboulay1974mechanism, heiss1999elucidating,
milhorat1993anatomical, stoodley2000pathophysiology, terae1994increased,
chang2003hypothesis, chang2004theoretical, greitz2006unraveling}, but
they provide widely different explanations on the mechanisms of syrinx
generation.  Most researchers, nevertheless, seem to agree on the following
points. First, the syrinx fluid is identical to the CSF, and there should
be some communication between the syrinx and the subarachnoid space. This
point is supported by many studies \cite{ellertsson1969syringomyelia,
ellertsson1969myelocystographic, li1987conventional, heiss2019origin},
although a few different opinions exist \cite{greitz2006unraveling,
koyanagi1997surgical}. Second, some derangement of CSF flow in the spinal
subarachnoid space causes syrinx both in Chiari-I malformation
\cite{wolpert1994chiari, bhadelia1995cerebrospinal, heiss1999elucidating,
hofmann2000phasecontrast, quigley2004cerebrospinal} and subarachnoid
arachnopathy \cite{klekamp1997treatment, brodbelt2003altered,
heiss2012pathophysiology, chang2014dorsal}. The cerebellar tonsils hamper
the CSF flow in the former and adhesive arachnoiditis in the latter.

@@

Some problem exists in the computer simulation of the motion of biological
fluids such as blood and CSF. Such bodily fluids move inside flexible
tubes. It, therefore, requires different analysis techniques from those
used in the engineering field, where the boundary of the conduit is
supposed to be solid. For this purpose, researchers widely used lumped
parameter models \cite{shi2011review, @kokalari2013review}. This model
considers the fluid flow inside a flexible tube in analogy to the flow of
electricity in an electric circuit. The accumulation of electricity in a
capacitor represents the expansion of a flexible tube and the accompanying
pressure elevation. An electrical resistor represents the frictional
resistance to flow. This model has a wide variety. For example, it may
model the whole cardiovascular system as one electric circuit, or it may
model it as a synthesis of multiple compartments of electric circuits
[@shi2011review; @kokalari2013review]. Our current model is the type of a
lumped parameter model with multiple compartments.  In this study, we did
not intend to make a quantitatively precise model of the spinal CSF flow
but to make a basic model that reveals the phenomenon underlying the
generation of syringomyelia. For this purpose, we adopted a revised version
of our previous lumped parameter model with multiple compartments.

@@

We analyzed the original normal system and two of its modifications.

@@

We present the system's responses as videos. Each video displays either all
or some of the following four values simultaneously: namely, the dural
tension (voltage in D capacitors in Figure 2), the canal tension (voltages
in C capacitors), the subarachnoid CSF flow (flows in R resistors), and the
canal flow (flows in r resistors). We plotted the node position on the
x-axis and the values at that node on the y-axis. The pressure values are
shown in $cmH_2O$, the flows in $ml/sec$. To plot different ranges of
values in a single plot, we multiplied the following values with certain
coefficients: namely, the canal tension with 50, the subarachnoid flow with
0.005, and the canal flow with $1.5\times10^{5}$. We plotted the
rightward flow in the positive and the leftward flow in the negative
direction.  Video 1 shows the original system's response to the sine wave
input on the cranial side. In this condition, CSF makes a smooth to-and-fro
movement in the subarachnoid space with the corresponding pressure wave
along the dura and the central canal. This response corresponds to the CSF
dynamics in a normal individual.

In Video 2, we increased the subarachnoid resistance $R_{25}$ by 20
times, thereby simulating a simple block of the subarachnoid flow in both
directions. In this condition, the dural tension showed a pressure drop at
node 25, corresponding to the increased resistance both in caudal and
rostral flow phases. This pressure drop caused an increase in canal flow in
a direction identical to the subarachnoid CSF flow. This increased canal
flow caused a transient increase in canal pressure distal to the block and
a decrease proximal to the block. However, these pressure changes
alternated according to the flow phase and did not produce a sustained
condition.

In Video 3, we replaced the subarachnoid resistor $R_{25}$
with a one-way valve that selectively resisted flow in the caudal direction
by 50 times. This time, the dural tension curve showed a pressure drop
across node 25, appearing only during the caudal-flow phase. Canal flow
showed alternating increase in caudal and rostral directions according to
the flow phase, which was similar to that in the simple block above.
However, there was a significant difference. Sustained high pressure
appeared in the central canal distal to the valve, and some sustained low
pressure in the segment proximal to the valve. The canal flow near node 25
markedly increased caudally in the caudal-flow phase and rostrally in the
rostral-flow phase. We took out the canal flow and showed it in Video 4.
Observing this video, we can see that the cumulative total of the caudal
flow is larger than that of the rostral flow, and it means that the CSF is
virtually pumped caudally at node 25.

@@

In this article, we theoretically analyzed the CSF movement in the spinal
cord using a lumped parameter model with multiple components. It simulated
a system with an elastic tube (dura) containing an elastic cylindrical
material (spinal cord) that itself contained a fluid channel (the central
canal). When we placed a direction-selective resistor in the subarachnoid
channel and evoked a to-and-fro pressure wave on this system, it produced a
sustained pressure elevation in the segment distal to the
direction-selective resistor. This phenomenon may explain the pathogenesis
of syringomyelia both in Chiari I malformation and syringomyelia associated
with arachnopathy.

Subarachnoid pressure affects the pressure inside the central canal. As
seen in Figure 2, the absolute pressure inside the central canal is the sum
of the subarachnoid and canal pressure (voltages of $C_k$ and $D_k$ in
electrical terms). Suppose a one-way valve selectively resists caudal flow
in the subarachnoid space at point A, and CSF makes a to-and-fro movement
across it. The caudal-flow phase creates a pressure drop across point A,
making the subarachnoid pressure in the distal segment smaller, and it
reduces the absolute canal pressure in the distal segment. Thus, the
pressure gradient in the central canal across point A increases, which
increases the distal CSF flow in the canal at that point.

On the contrary, the rostral flow does not create a pressure drop. Although
some of the CSF pumped caudally will flow back rostrally, its amount will
be smaller. The net result will be that some CSF is pumped caudally in one
cycle of the to-and-fro movement. Thus, CSF gradually accumulates in the
distal segment of the resistance. We hypothesize that this is the mechanism
underlying the syrinx generation.

@@

\begin{center}
\includegraphics[width=\textwidth]{pumping_mechanism_close.jpg}
\includegraphics[width=\textwidth]{pumping_mechanism_open.jpg}
\end{center}

>>>>

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{newfloat}
\DeclareFloatingEnvironment{video}
\title{Direction-Selective Resistance to Cerebrospinal-Fluid Flow As
the Mechanism of Syrinx Generation in Syringomyelia}

@@

\keywords{syringomyelia, pathophysiology, simulation}
\input{embed_video.tex}
\begin{abstract}
    The pathophysiology of syringomyelia is not well understood. The main
    theoretical problem is how cerebrospinal fluid (CSF) enters from
    low-pressure subarachnoid space to high-pressure syrinx and remains
    inside the syrinx. We approached this problem with computer simulation
    of the CSF flow in the spine.
\end{abstract}
\begin{document}

@@

@@

The pathophysiology of syringomyelia is still poorly understood. A number
of hypotheses exist in the literature \cite{gardner1958mechanism,
williams1980pathogenesis, milhorat1999chiari, ball1972pathogenesis,
klekamp2002pathophysiology, duboulay1974mechanism, heiss1999elucidating,
milhorat1993anatomical, stoodley2000pathophysiology, terae1994increased,
chang2003hypothesis, chang2004theoretical, greitz2006unraveling}, but they
provide widely different explanations on the mechanisms of syrinx
generation.  Nevertheless, most researchers seem to agree on the following
points. First, the syrinx fluid is identical to the CSF, and there should
be some communication between the syrinx and the subarachnoid space. This
point is supported by many studies \cite{ellertsson1969syringomyelia,
ellertsson1969myelocystographic, li1987conventional, heiss2019origin},
although a few different opinions exist \cite{greitz2006unraveling,
koyanagi1997surgical}. Second, some derangement of CSF flow in the spinal
subarachnoid space causes syrinx both in Chiari-I malformation
\cite{wolpert1994chiari, bhadelia1995cerebrospinal, heiss1999elucidating,
hofmann2000phasecontrast, quigley2004cerebrospinal} and subarachnoid
arachnopathy \cite{klekamp1997treatment, brodbelt2003altered,
heiss2012pathophysiology, chang2014dorsal}. Notably, the cerebellar tonsil
deranges the CSF flow in the former and adhesive arachnoiditis in the
latter.

@@

Some problem exists in the computer simulation of the motion of biological
fluids such as blood and CSF. Such bodily fluids move inside flexible
tubes. It, therefore, requires different analysis techniques from those
used in the engineering field, where the boundary of the conduit is
supposed to be solid. For this purpose, researchers widely used lumped
parameter models \cite{shi2011review, kokalari2013review}. This model
considers the fluid flow inside a flexible tube in analogy to the flow of
electricity in an electric circuit. The accumulation of electricity in a
capacitor represents the expansion of a flexible tube and the accompanying
pressure elevation. An electrical resistor represents the frictional
resistance to flow. This model has a wide variety. For example, it may
model the whole cardiovascular system as one electric circuit, or it may
model it as a synthesis of multiple compartments of electric circuits
\cite{shi2011review, kokalari2013review}. Our current model is the type of a
lumped parameter model with multiple compartments.  In this study, we did
not intend to make a quantitatively precise model of the spinal CSF flow
but to make a basic model that reveals the phenomenon underlying the
generation of syringomyelia. For this purpose, we adopted a revised version
of our previous lumped parameter model with multiple compartments.

@@

We analyzed the original normal system and two of its modifications. In the
first modification, we simply increased the subarachnoid resistance at
point 25 ($R_{25}$) by 20 times. In the second modification, we placed a
direction-selective subarachnoid resistance at point 25, so that only the
resistance to the caudal flow would be increased by 20 times.

@@

We present the systems' responses to twenty cycles of to-and-fro waves as
animations. The x-axis of each animation represents the 100 nodes of the
circuit laid out from the cranial to caudal direction. The y-axis displays
some of the following four measures: the dural tension (voltage in D
capacitors in Figure 2), the canal tension (voltages in C capacitors), the
subarachnoid CSF flow (flows in R resistors), and the canal flow (flows in
r resistors). We plotted the caudal flow in the positive and the rostral
flow in the negative direction. The pressure values are shown in $cmH_2O$,
the flows in $ml/sec$. To plot the four measures in a single animation, we
multiplied the following three measures---the canal tension, the
subarachnoid CSF flow, and the canal flow---with the following
coefficients, respectively: namely, the canal tension with 50, the
subarachnoid flow with 0.005, and the canal flow with $1.5\times10^{5}$.

Video \ref{video:norm} shows the original system's response representing
the normal condition. In this state, CSF makes a smooth to-and-fro movement
in the subarachnoid space with the corresponding pressure wave along the
dura and the central canal. CSF also makes to-and-fro movements in the
canal.

\begin{video}[hbt]
    \embedvideo*{\includegraphics[width=\textwidth]{thumbnail_norm.jpg}}{totalAnimationNorm.mp4}
    \caption{Video showing the normal circuit response}
    \label{video:norm}
\end{video}

In Video \ref{video:simple_block}, we increased the subarachnoid resistance
$R_{25}$ by 20 times, thereby simulating a simple block of the subarachnoid
flow in both directions. In this condition, both the caudal and rostral
flow along the resistance produced a drop of the dural tension downstream
to the resistance, corresponding to the increased resistance both in the
caudal and rostral flow phases. This pressure drop caused an increase in
canal flow in the same direction as the subarachnoid flow. This increased
canal flow caused a transient increase in the canal pressure distal to the
block and a decrease proximal to the block. However, these pressure changes
alternated along the alternation of the flow direction and did not produce
a sustained pressure increase.

\begin{video}[hbt]
    \embedvideo*{\includegraphics[width=\textwidth]{thumbnail_simple_block.jpg}}{totalAnimationSb.mp4}
    \caption{Video showing the response of the circuit with a simple
    resistor at point 25}
    \label{video:simple_block}
\end{video}

\begin{video}[hbt]
    \embedvideo*{\includegraphics[width=\textwidth]{thumbnail_oneway.jpg}}{totalAnimation.mp4}
    \caption{Video showing the response of the circuit with a
    direction-selective resistor at point 25}
    \label{video:oneway}
\end{video}

\begin{video}[hbt]
    \embedvideo*{\includegraphics[width=\textwidth]{thumbnail_oneway.jpg}}{canalFlowOneway.mp4}
    \caption{Video showing the canal flow in the model with
    direction-selective resistance at point 25}
    \label{video:canal_flow_oneway}
\end{video}

\begin{video}[hbt]
    \embedvideo*{\includegraphics[width=\textwidth]{thumbnail_oneway.jpg}}{canalFlowComparison.mp4}
    \caption{Video showing the canal flow in the two models: simple
    resistance at point 25 and direction-selective resistance at point 25}
    \label{video:canal_comparison}
\end{video}

In Video \ref{video:oneway}, we replaced the subarachnoid resistor $R_{25}$
with a direction-selective resistor whose resistance to the rostral flow
was unchanged but that to the caudal flow was increased by 20 times.  This
time, as shown in Video \ref{video:oneway}, sustained high pressure
appeared in the central canal in the segment distal to the replaced
resistor, and sustained low pressure in the segment proximal to it.  This
sustained pressure gradually accumulated as the flow cycle proceeded.  The
dural tension showed a pressure drop at node 25 only during the caudal-flow
phase.  The to-and-fro canal flow increased near the node 25 similarly to
that in the simple block above, but, this time, the increase was larger in
the caudal direction.

We took out the canal flow and showed it in Video 4.  Observing this video,
we can see that the cumulative total of the caudal flow is larger than that
of the rostral flow, and it means that the CSF is virtually pumped caudally
at node 25. To further elucidate this point, we plotted the canal flow in
the simple block as shown in Video \ref{video:simple_block} and that in the
one-way block as shown in Video \ref{video:oneway}. We can clearly see in
this video that the caudal flow in the canal is

@@

In this article, we theoretically analyzed the CSF movement in the spinal
cord using a lumped parameter model with multiple components. It simulated
a system with an elastic tube (dura) containing an elastic cylindrical
material (spinal cord) that itself had a fluid channel inside (the central
canal). When we placed a direction-selective resistor in the subarachnoid
space and evoked a to-and-fro pressure wave on this system, it produced a
sustained pressure elevation in the segment distal to the
direction-selective resistor. This phenomenon may explain the pathogenesis
of syringomyelia both in Chiari I malformation and syringomyelia associated
with arachnopathy.

Subarachnoid pressure affects the pressure inside the central canal. As
seen in Figure \ref{fig:circuit}, the absolute pressure inside the central
canal is the sum of the subarachnoid and canal pressure (voltages of $C_k$
and $D_k$ in electrical terms). Suppose a one-way valve selectively resists
caudal flow in the subarachnoid space at point A. Caudal CSF flow creates a
pressure drop across this valve, with the distal pressure smaller than the
proximal one. It decreases the absolute canal pressure distal to point A
because it is the sum of the subarachnoid pressure and the canal pressure.
It, therefore, creates a pressure gradient in the central canal across
point A, thereby increasing the distal CSF flow in the canal at that point
(Figure \ref{fig:pump_close}).

\begin{figure}[hbt]
    \centering
    \includegraphics[width=\textwidth]{pumping_mechanism_close.jpg}
    \caption{Increased central-canal flow across a direction-selective
    resistance in the subarachnoid space}
    \label{fig:pump_close}
\end{figure}

On the contrary, reverse flow, not encountering resistance, does not create
a pressure drop (Figure \ref{fig:pump_open}). Although some of the CSF that
had been pumped caudally during the caudal-flow phase will flow back
rostrally, its amount will be smaller. The net result will be that some CSF
is pumped caudally in one cycle of the to-and-fro movement. Thus, CSF
gradually accumulates in the distal segment of the resistance (Video 3). We
hypothesize that this is the mechanism underlying the syrinx generation.

\begin{figure}[hbt]
    \centering
    \includegraphics[width=\textwidth]{pumping_mechanism_open.jpg}
    \caption{Normal central-canal flow across a direction-selective
    resistance during the reverse flow}
    \label{fig:pump_open}
\end{figure}

@@
